### 2016-01-27 doc/lda.Rmd

---
title: "LDA exploration"
author: "Luke W. Johnston"
date: "January 19, 2016"
output: 
    rmarkdown::html_vignette:
        toc: true
---

# Load up packages and data

```{r setup, collapse=TRUE}
knitr::opts_knit$set(root.dir = '../')
knitr::opts_chunk$set(collapse = TRUE, fig.height = 8, fig.width = 10)
```

```{r loadData, message=FALSE}
source('.Rprofile')
suppressMessages(run_setup())
load_data()
dim(ds)
```

# Prepare the data for the LDA

```{r ldaData}
lda_ds <- ds %>%
    filter(VN == 0) %>%
    mutate(BCF = rstatsToolkit::aide.tertile(lISSI2) %>%
               factor(., labels = c('1st tertile', '2nd tertile', '3rd tertile'), ordered = TRUE)) %>%
    select(BCF, matches('^ne\\d\\d')) %>%
    mutate_each(funs(as.numeric(scale(.))), -BCF) %>%
    na.omit() %>% 
    arrange(BCF)

# include also the 3-yr and 6-yr data for training/testing.

#fit <- qda(BCF ~ ., data = lda_ds)
fit <- lda(BCF ~ ., data = lda_ds)
fit.cv <- lda(BCF ~ ., data = lda_ds, CV = TRUE)
prop_lda <- round(fit$svd^2/sum(fit$svd^2) * 100, 2)

# proportion explained by each LD
prop_lda
```

```{r testingLDA}
fit.class <- predict(fit, newdata = lda_ds[, -1])$class %>% 
    factor(., ordered = TRUE)
table(fit.class, lda_ds[[1]])
table(fit.cv$class, lda_ds[[1]])
```

```{r FAperBCF}
p_data <- fit$means %>%
    as.data.frame() %>%
    add_rownames('Group') %>%
    gather(FA, Mean,-Group) %>%
    mutate(Group = Group %>% factor(., c('1st tertile', '2nd tertile', '3rd tertile'), ordered = TRUE)) %>%
    arrange(Group, Mean) %>%
    mutate(FA = paste0(Group, FA) %>%
               factor(., unique(.)))
FA.orig <- p_data$FA %>% 
    as.character()
FA.label <- p_data$FA %>%
    gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
    gsub('n(\\d)', 'n-\\1', .) %>%
    as.character()
p_data %>%
    ggplot(aes(x = Mean, y = FA)) +
    geom_point() +
    geom_vline(xintercept = 0) +
    facet_wrap(~ Group, scale = 'free_y') +
    scale_y_discrete(breaks = FA.orig, label = FA.label) +
    ggthemes::theme_tufte(12, 'sans') +
    theme(panel.margin = unit(2, 'lines'),
          panel.grid = element_line(colour = 'black'),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_blank())

```

```{r FAperLD}
p_data2 <- fit$scaling %>%
    as.data.frame() %>%
    add_rownames('Var') %>%
    gather(LD, Value, -Var) %>%
    arrange(LD, Value) %>% 
    mutate(Var = factor(paste0(LD, Var), unique(paste0(LD, Var))))
FA.orig <- p_data2$Var %>% 
    as.character()
FA.label <- p_data2$Var %>%
    gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
    gsub('n(\\d)', 'n-\\1', .) %>%
    as.character()
p_data2 %>% 
    ggplot(aes(x = Value, y = Var)) +
    geom_point() +
    geom_vline(xintercept = 0) +
    facet_wrap(~LD, scale = 'free_y') +
    scale_y_discrete(breaks = FA.orig, label = FA.label) +
    ggthemes::theme_tufte(12, 'sans') +
    theme(panel.margin = unit(2, 'lines'),
          panel.grid = element_line(colour = 'black'),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_blank())
```

```{r LDbyBCF}
p_fit <- predict(object = fit, newdata = lda_ds)
# predict on follow up visits too.
p_ds <- data.frame(BCF = lda_ds[,"BCF"], lda = p_fit$x)
ggplot(p_ds, aes(lda.LD1, lda.LD2, group = BCF)) +
    geom_vline(xintercept = 0, linetype = 'dotted') +
    geom_hline(yintercept = 0, linetype = 'dotted') +
    stat_density2d(aes(alpha = ..level.., colour = BCF), size = 2) +
    labs(x = paste0("LD1 (", prop_lda[1], "%)"),
         y = paste0('LD2 (', prop_lda[2], '%)')) +
    ggthemes::theme_tufte(12, 'sans') +
    scale_alpha(guide = 'none') +
    theme(legend.title = element_blank(),
          legend.position = c(1,0),
          legend.justification = c(1,0))
```