### 2016-02-23 doc/lda.Rmd
Changed lda analysis over to lcmm + plsda

---
title: "LCMM, PLS-DA, and LDA exploration"
author: "Luke W. Johnston"
date: "January 19, 2016"
output: 
    rmarkdown::html_vignette:
        toc: true
---

# Load up packages and data

```{r setup, collapse=TRUE}
knitr::opts_knit$set(root.dir = '../')
knitr::opts_chunk$set(collapse = TRUE, fig.height = 8, fig.width = 10)
```

```{r loadData, message=FALSE}
source('.Rprofile')
suppressMessages(run_setup())
load_data()
dim(ds)
```

# Latent class mixed model

```{r lccmPkg}
ds.lcmm <- ds %>% 
    select(SID, lISSI2, VN) %>% 
    na.omit() %>% 
    arrange(SID, VN)
fit <- lcmm(lISSI2 ~ VN,
            mixture = ~VN, random = ~VN, subject = 'SID',
            ng = 3, idiag = TRUE, link = '3-equi-splines', 
            data = ds.lcmm)
as.data.frame(table(fit$pprob$class)) %>% 
    rename(Group = Var1, N = Freq)

ds.lcmm2 <- 
    full_join(ds.lcmm,
              as.data.frame(fit$pprob[,1:2]),
              by = 'SID') %>%
    mutate(class = factor(class, labels = c('Mid', 'High', 'Low')))
ggplot(ds.lcmm2, aes(VN, lISSI2, group = SID, colour = class)) +
    geom_line(aes(group = SID, colour = class), size = 0.1, alpha = 0.9) +
    geom_smooth(aes(group = class), method = "loess", size = 2)  +
    labs(x = "Time", y = "ISSI-2", colour = "Latent Class") +
    theme_minimal()
ds <- full_join(ds, ds.lcmm2)
dim(ds)
```

# Partial least squares discriminant analysis

```{r caretPkg}
ds.plsda <- ds %>%
    filter(VN == 0) %>% 
    select(class, matches('^ne\\d\\d')) %>% 
    na.omit()

if (length(nearZeroVar(ds.plsda[, 2:23])) != 0) message('Near zero variance, check into it')
ds.plsda2 <- predict(preProcess(ds.plsda), ds.plsda)
y <- as.factor(ds.plsda2[[1]])
x <- as.matrix(ds.plsda2[2:23])
fit <- plsda(x, y, probMethod = 'Bayes')

mismatch <- confusionMatrix(predict(fit, x), y) %>% 
    .$table %>% 
    as.data.frame %>% 
    mutate(Match = ifelse(Prediction == Reference, 'yes', 'no')) %>% 
    group_by(Match) %>% 
    summarise(N = sum(Freq)) %>% 
    ungroup() %>% 
    mutate(Total = sum(N),
           Percent = round((N / Total) * 100, 1))
mismatch

plot(fit, plottype = 'loadings')
plot(fit, plottype = 'correlation')

exp.prop <- round((fit$Xvar/fit$Xtotvar)*100, 2)
exp.prop
fit$loadings %>% 
    as.data.frame.matrix %>% 
    add_rownames %>% 
    rename(Comp1 = `Comp 1`, Comp2 = `Comp 2`) %>% 
    mutate(rowname = rowname %>% 
               renaming_fa() %>% 
               factor(., levels = unique(.))) %>% 
    ggplot(aes(Comp1, Comp2)) +
    geom_point(size = 0.5) +
    geom_text(aes(label = rowname), hjust = 0, vjust = 0) +
    labs(x = paste0('Component 1 (', round(exp.prop[1], 1), '%)'), 
         y = paste0('Component 2 (', round(exp.prop[2], 1), '%)')) +
    ggthemes::theme_tufte(11, 'sans') +
    theme(panel.grid = element_line(),
          panel.grid.minor = element_blank())

fit$scores %>% 
    as.data.frame.matrix %>% 
    bind_cols(ds.plsda['class']) %>% 
    rename(Comp1 = `Comp 1`, Comp2 = `Comp 2`) %>% 
    ggplot(aes(Comp1, Comp2, colour = class)) +
    stat_density2d(ggplot2::aes(alpha = ..level.., colour = class), size = 2) +
    scale_alpha(guide = 'none') +
    scale_color_discrete('LCMM Class') +
    labs(x = paste0('Component 1 (', round(exp.prop[1], 1), '%)'), 
         y = paste0('Component 2 (', round(exp.prop[2], 1), '%)')) +
    ggthemes::theme_tufte(11, 'sans') +
    theme(panel.grid = element_line(),
          panel.grid.minor = element_blank())

```

# Linear discriminant analysis

```{r ldaData}
ds.prep <- ds %>%
    filter(VN == 0) %>%
    select(class, matches('^ne\\d\\d')) %>% 
    mutate_each(funs(as.numeric(scale(.))), -class) %>%
    rename(Class = class)

df_y1 <- ds.prep %>% 
    na.omit()

# include also the 3-yr and 6-yr data for training/testing.
fit_y1 <- MASS::lda(Class ~ ., data = df_y1)
```

```{r testingLDA}
fit.class_y1 <- MASS:::predict.lda(fit_y1, newdata = df_y1[, -1])$class %>% 
    factor(., ordered = TRUE)

table(fit.class_y1, df_y1[[1]])

# Amount of participants who were misclassified
table(fit.class_y1, df_y1[[1]]) %>% 
    as.data.frame() %>% 
    filter(fit.class_y1 != Var2) %>% 
    summarize(count = sum(Freq))

# Amount of participants who were correctly classified
# table(fit.class_y1, df_y1[[1]]) %>% 
#     as.data.frame() %>% 
#     filter(fit.class_y1 == Var2) %>% 
#     fit.class_y1 <- predict(fit_y1, newdata = df_y1[, -1])$class %>% 
#     factor(., ordered = TRUE)summarize(count = sum(Freq))

```

```{r FAperBCF}
seer::visualize(
    fit_y1, 'means', label.rename.fun = function(x)
        x %>%
        gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
        gsub('n(\\d)', 'n-\\1', .)
    ) %>% 
    seer::vision_simple()
```

```{r FAperLD}
seer::visualize(
    fit_y1, 'scaling', label.rename.fun = function(x)
        x %>%
        gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
        gsub('n(\\d)', 'n-\\1', .)
    ) %>% 
    seer::vision_simple()
```

```{r LDbyBCF}
seer::visualize(fit_y1, orig.data = df_y1, group.name = 'Class') %>% 
    seer::vision_simple()
```