### 2016-07-06 doc/gee.Rmd
First commit for submission code

---
title: "GEE models"
author: "Luke W. Johnston"
output: 
    rmarkdown::html_vignette:
        toc: true
---

# Set up and load dataset

```{r setup, message = TRUE, collapse=TRUE}
knitr::opts_knit$set(root.dir = '../')
knitr::opts_chunk$set(collapse = TRUE)
```

```{r}
source('.Rprofile')
suppressMessages(run_setup())
ds <- load_data()
ds %>% 
    select(VN, TAG) %>% 
    na.omit() %>% 
    group_by(VN) %>% 
    summarize(n = n())

gee_table <- function(data, caption, p.val = 0.05) {
    data %>%
        filter(p.value < p.val) %>%
        select(Yterms, Xterms, term, unit, estimate, p.value) %>%
        mutate_each(funs(round(., 2)), estimate, p.value) %>%
        arrange(unit, Yterms) %>% 
        pander(caption)
}
```

# Prep and wrangle data into GEE form

```{r geeData}
gee_df <- get_gee_data(ds)

outcomes <- c('linvHOMA', 'lISI', 'lIGIIR', 'lISSI2')
outcomes_is <- c('linvHOMA', 'lISI')
outcomes_bcf <- c('lIGIIR', 'lISSI2')
ne_pct <- grep('^pct_ne', names(gee_df), value = TRUE)
ne_conc <- c('TotalNE', grep('^ne\\d\\d', names(gee_df), value = TRUE))
covariates0 <- c('VN')
covariates_bcf <- c('VN', 'Waist', 'Sex', 'Ethnicity', 'BaseAge', 'ALT', 'FamHistDiab')
covariates <- covariates_is <- covariates_bcf
covariates_is_tag <- c(covariates_is, 'BaseTAG')
covariates_bcf_tag <- c(covariates_bcf, 'BaseTAG')
```

# Unadjusted GEE model

```{r geeResults_Unadj}
gee_results0 <- dplyr::bind_rows(
    analyze_gee(gee_df, outcomes, ne_pct, covariates0, 'mol%', adj.p = TRUE),
    analyze_gee(gee_df, outcomes, ne_conc, covariates0, 'nmol/mL', adj.p = TRUE)
    ) %>% 
    tidy_gee_results()
gee_table(gee_results0, 'Unadjusted')
```

# Adjusted with or without TAG for IS

```{r geeResults_IS}
gee_results_is <- bind_rows(
    analyze_gee(gee_df, outcomes_is, ne_pct, covariates_is, 'mol%', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_conc, covariates_is, 'nmol/mL', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_pct, covariates_is_tag, 'mol%', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_conc, covariates_is_tag, 'nmol/mL', adj.p = TRUE)
    ) %>% 
    tidy_gee_results()

gee_table(gee_results_is, 'Without or with TAG for IS')
gee_results_is %>% 
    select(Yterms, Xterms, sample.total) %>% 
    group_by(Yterms) %>% 
    summarise(MinSample = min(sample.total),
              MaxSample = max(sample.total)) %>% 
    pander('Sample size for IS')
```

# Adjusted with or without TAG for BCF

```{r geeResults_BCF}
gee_results_bcf <- bind_rows(
    analyze_gee(gee_df, outcomes_bcf, ne_pct, covariates_bcf, 'mol%', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_bcf, ne_conc, covariates_bcf, 'nmol/mL', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_bcf, ne_pct, covariates_bcf_tag, 'mol% (w/ tag)', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_bcf, ne_conc, covariates_bcf_tag, 'nmol/mL (w/ tag)', adj.p = TRUE)
    ) %>% 
    tidy_gee_results()

gee_table(gee_results_bcf, 'Without or with TAG for BCF')
gee_results_bcf %>% 
    select(Yterms, Xterms, sample.total) %>% 
    group_by(Yterms) %>% 
    summarise(MinSample = min(sample.total),
              MaxSample = max(sample.total)) %>% 
    pander('Sample size for BCF')
```

# Adjusting using a single model for both IS and BCF (cancelled)

Decided on using only one model. Not running the below code but keeping it for 
now for record-keeping.

```{r geeResults_SingleModel, eval = FALSE}
gee_results <- bind_rows(
    analyze_gee(gee_df, outcomes, ne_pct, covariates_is, 'mol% (IS model)', adj.p = TRUE),
    analyze_gee(gee_df, outcomes, ne_conc, covariates_is, 'nmol/mL (IS model)', adj.p = TRUE),
    analyze_gee(gee_df, outcomes, ne_pct, covariates_bcf, 'mol% (BCF model)', adj.p = TRUE),
    analyze_gee(gee_df, outcomes, ne_conc, covariates_bcf, 'nmol/mL (BCF model)', adj.p = TRUE)
    ) %>% 
    tidy_gee_results()

gee_table(gee_results, 'Comparing IS vs BCF models.')
```

# Identifying which variable changes association for IS

```{r geeResults_VariableInfluence}
covar_base <- c('VN', 'Sex', 'Ethnicity', 'BaseAge')
covar_wc <- c(covar_base, 'Waist')
covar_alt <- c(covar_base, 'ALT')
covar_fh <- c(covar_base, 'FamHistDiab')
gee_results <- bind_rows(
    analyze_gee(gee_df, outcomes_is, ne_pct, covar_base, 'mol% (base)', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_conc, covar_base, 'nmol/mL (base)', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_pct, covar_wc, 'mol% (wc)', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_conc, covar_wc, 'nmol/mL (wc)', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_pct, covar_fh, 'mol% (fh)', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_conc, covar_fh, 'nmol/mL (fh)', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_pct, covar_alt, 'mol% (alt)', adj.p = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_conc, covar_alt, 'nmol/mL (alt)', adj.p = TRUE)
    ) %>% 
    tidy_gee_results()

gee_table(gee_results, 'Covariate influencing results.')
```

# Grouping NEFA into common classes (SFA, MUFA, PUFA)

```{r geeResults_FAGroupings}
sfa <- grep('0$', c(ne_conc), value = TRUE)
mufa <- grep('1n.*$', c(ne_conc), value = TRUE)
pufa <- grep('[2-6]n.*$', c(ne_conc), value = TRUE)
n3pufa <- grep('[2-6]n3$', c(ne_conc), value = TRUE)
n6pufa <- grep('[2-6]n6$', c(ne_conc), value = TRUE)

gee_df2 <- ds %>% 
    mutate(
        SFA = rowSums(.[sfa]),
        pct_SFA = (SFA / TotalNE) * 100,
        MUFA = rowSums(.[mufa]),
        pct_MUFA = (MUFA / TotalNE) * 100,
        PUFA = rowSums(.[pufa]),
        pct_PUFA = (PUFA / TotalNE) * 100,
        n3PUFA = rowSums(.[n3pufa]),
        pct_n3PUFA = (n3PUFA / TotalNE) * 100,
        n6PUFA = rowSums(.[n6pufa]),
        pct_n6PUFA = (n6PUFA / TotalNE) * 100
        ) %>% 
    get_gee_data(fa = '(^|^pct_).*(SFA|MUFA|PUFA)')

gee_results <- bind_rows(
    analyze_gee(gee_df2, outcomes, c('SFA', 'MUFA', 'PUFA', 'n3PUFA', 'n6PUFA'), 
                covariates, 'nmol/mL', adj.p = TRUE), 
    analyze_gee(gee_df2, outcomes, c('pct_SFA', 'pct_MUFA', 'pct_PUFA', 
                                     'pct_n3PUFA', 'pct_n6PUFA'), 
                covariates, 'mol%', adj.p = TRUE)
    ) %>% 
    tidy_gee_results()

gee_table(gee_results, 'Results using fatty acid classes (saturated, etc).')
```

# Interaction for IS

```{r geeResults_IS_int}
gee_results_is_int <- bind_rows(
    analyze_gee(gee_df, outcomes_is, ne_pct, covariates_is, 'mol%', adj.p = TRUE, int = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_conc, covariates_is, 'nmol/mL', adj.p = TRUE, int = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_pct, covariates_is_tag, 'mol% (w/ tag)', adj.p = TRUE, int = TRUE),
    analyze_gee(gee_df, outcomes_is, ne_conc, covariates_is_tag, 'nmol/mL (w/ tag)', adj.p = TRUE, int = TRUE)
    ) %>% 
    tidy_gee_results()
gee_table(gee_results_is_int, 'Interaction for time for IS')
```

```{r geeResults_IS_int_sex}
gee_results_is_int_sex <- bind_rows(
    analyze_gee(gee_df, outcomes_is, ne_pct, covariates_is, 'mol%', adj.p = TRUE, int = TRUE, intvar = 'Sex'),
    analyze_gee(gee_df, outcomes_is, ne_conc, covariates_is, 'nmol/mL', adj.p = TRUE, int = TRUE, intvar = 'Sex'),
    analyze_gee(gee_df, outcomes_is, ne_pct, covariates_is_tag, 'mol% (w/ tag)', adj.p = TRUE, int = TRUE, intvar = 'Sex'),
    analyze_gee(gee_df, outcomes_is, ne_conc, covariates_is_tag, 'nmol/mL (w/ tag)', adj.p = TRUE, int = TRUE, intvar = 'Sex')
    ) %>% 
    tidy_gee_results()

gee_table(gee_results_is_int_sex, 'Interactions between male and female for IS.')
```

```{r geeResults_IS_int_ethn}
gee_results_is_int_ethn <- bind_rows(
    analyze_gee(gee_df, outcomes_is, ne_pct, covariates_is, 'mol%', 
                adj.p = TRUE, int = TRUE, intvar = 'Ethnicity'),
    analyze_gee(gee_df, outcomes_is, ne_conc, covariates_is, 'nmol/mL', 
                adj.p = TRUE, int = TRUE, intvar = 'Ethnicity'),
    analyze_gee(gee_df, outcomes_is, ne_pct, covariates_is_tag, 'mol% (w/ tag)', 
                adj.p = TRUE, int = TRUE, intvar = 'Ethnicity'),
    analyze_gee(gee_df, outcomes_is, ne_conc, covariates_is_tag, 'nmol/mL (w/ tag)', 
                adj.p = TRUE, int = TRUE, intvar = 'Ethnicity')
    ) %>% 
    tidy_gee_results()

gee_table(gee_results_is_int_ethn, 'Interactions between ethnicities for IS.')
```

# Interaction for BCF

```{r geeResults_BCF_int}
gee_results_bcf_int <- bind_rows(
    analyze_gee(gee_df, outcomes_bcf, ne_pct, covariates_bcf, 'mol%', adj.p = TRUE, int = TRUE),
    analyze_gee(gee_df, outcomes_bcf, ne_conc, covariates_bcf, 'nmol/mL', adj.p = TRUE, int = TRUE),
    analyze_gee(gee_df, outcomes_bcf, ne_pct, covariates_bcf_tag, 'mol% (w/ tag)', adj.p = TRUE, int = TRUE),
    analyze_gee(gee_df, outcomes_bcf, ne_conc, covariates_bcf_tag, 'nmol/mL (w/ tag)', adj.p = TRUE, int = TRUE)
    ) %>% 
    tidy_gee_results()

gee_table(gee_results_bcf_int, 'Interactions for time for BCF.')
```

```{r geeResults_BCF_int_sex}
gee_results_bcf_int_sex <- bind_rows(
    analyze_gee(gee_df, outcomes_bcf, ne_pct, covariates_bcf, 'mol%', 
                adj.p = TRUE, int = TRUE, intvar = 'Sex'),
    analyze_gee(gee_df, outcomes_bcf, ne_conc, covariates_bcf, 'nmol/mL', 
                adj.p = TRUE, int = TRUE, intvar = 'Sex'),
    analyze_gee(gee_df, outcomes_bcf, ne_pct, covariates_bcf_tag, 'mol% (w/ tag)', 
                adj.p = TRUE, int = TRUE, intvar = 'Sex'),
    analyze_gee(gee_df, outcomes_bcf, ne_conc, covariates_bcf_tag, 'nmol/mL (w/ tag)', 
                adj.p = TRUE, int = TRUE, intvar = 'Sex')
    ) %>% 
    tidy_gee_results()

gee_table(gee_results_bcf_int_sex, 'Interactions between male and female for BCF.')
```

```{r geeResults_BCF_int_ethn}
gee_results_bcf_int_ethn <- bind_rows(
    analyze_gee(gee_df, outcomes_bcf, ne_pct, covariates_bcf, 'mol%', 
                adj.p = TRUE, int = TRUE, intvar = 'Ethnicity'),
    analyze_gee(gee_df, outcomes_bcf, ne_conc, covariates_bcf, 'nmol/mL', 
                adj.p = TRUE, int = TRUE, intvar = 'Ethnicity'),
    analyze_gee(gee_df, outcomes_bcf, ne_pct, covariates_bcf_tag, 'mol% (w/ tag)', 
                adj.p = TRUE, int = TRUE, intvar = 'Ethnicity'),
    analyze_gee(gee_df, outcomes_bcf, ne_conc, covariates_bcf_tag, 'nmol/mL (w/ tag)', 
                adj.p = TRUE, int = TRUE, intvar = 'Ethnicity')
    ) %>% 
    tidy_gee_results()

gee_table(gee_results_bcf_int_ethn, 'Interactions between ethnicities for BCF.')
```