### 2017-07-02 doc/esm-methods.Rmd
added esm-methods file

---
output: fost::diabetologia_manuscript
---

```{r setup, include=FALSE}
devtools::load_all()
set_options()
knitr::opts_chunk$set(results = "asis")
```

# ESM Methods

## Equations for outcomes

The equations for calculating the outcome variables are listed below. HOMA2-%S was 
calculated using the Excel found at https://www.dtu.ox.ac.uk/homacalculator/.

```{r}
PROMISE.methods::insert_formula("ISI")
```

```{r}
PROMISE.methods::insert_formula("HOMA")
```

```{r}
PROMISE.methods::insert_formula("IGIIR")
```

```{r}
PROMISE.methods::insert_formula("ISSI2")
```

Area under the curve (AUC) for both glucose (G) and insulin (I) was computed
using the trapezoidal rule. ISI~OGTT~ (ISI) uses all the glucose and insulin
values from the oral glucose tolerance (the overall mean).

## Further description of statistical methods

Generalized estimating equations (GEE) is analogous to mixed effects models, 
except it emphasizes population-level trends over individual-level trends that 
mixed models use. Because it can handle autocorrelations in the outcome (as 
would be seen in variables measured over time), GEE modeling is well suited to
longitudinal data. Latent class mixed models (LCMM) is a form of mixed effects
modeling that tries to categorize individuals into groups that have similar
clusters; in the case of the present analysis are the trajectories in beta-cell
function over time. Because mixed models use the participant as the random
effect, LCMM can identify individuals who similarly change over time.

Partial least squares (PLS) regression is a technique similar to principal
component analysis (PCA) that is designed to extract meaningful information from
multivariate, high dimensionality data (e.g. as in metabolomic and other -omic type 
analyses). These types of methods try to extract as much of the variance in the 
data into a smaller number of components or factors. The difference between PCA 
and PLS is that PLS is a supervised method, while PCA is not. This means that
PLS uses a response variable(s), i.e. an outcome or **Y**, to describe the
variation in the predictors (**X**) while PCA only describes the variation 
inherent in **X**. Because of this, PLS can be better at predicting the 
contributions of predictor variables against an outcome variable.