### 2017-07-03 doc/esm-figures.Rmd
esm figures file done

---
output: fost::diabetologia_manuscript
---

```{r setup, include=FALSE}
devtools::load_all()
set_options()

# Include captions below using `captioner` package via library(captioner)
sfig <- captioner::captioner(prefix = 'ESM Fig.')
cite_sf <- pryr::partial(sfig, display = 'cite')

sfig_consort <- sfig('consort', 'CONSORT diagram of sample size at each examination visit.')
sfig_dag <- sfig("dagitty", "Directed acyclic graph taken from DAGitty (DAGitty.net).")
sfig_corr_conc <- sfig('corr_conc', 'Pearson correlation heatmap of non-esterified fatty acids (*nmol/mL*) and basic PROMISE participant characteristics for the baseline visit (2004-2006). Darkness of the colour indicates the magnitude of the correlation, with orange indicating positive and blue indicating negative correlations.')
sfig_corr_pct <- sfig('corr_pct', 'Pearson correlation heatmap of non-esterified fatty acids (*mol%*) and basic PROMISE participant characteristics for the baseline visit (2004-2006). Darkness of the colour indicates the magnitude of the correlation, with orange indicating positive and blue indicating negative correlations.')
```

# ESM Figures

![`r sfig_consort`](../img/flowDiagramSample.png)

![`r sfig_dag`](../img/dagitty.png)

```{r ESM_Fig3, fig.cap=sfig_corr_conc}
plot_corr_heatmap(corr_results_conc, unit = "nmol/mL")
```

```{r ESM_Fig4, fig.cap=sfig_corr_pct}
plot_corr_heatmap(corr_results_pct, unit = "mol%")
```