### 2017-07-03 doc/esm-tables.Rmd
moved esm tables into own file and updated

---
output: fost::diabetologia_manuscript
---

```{r setup, include=FALSE}
devtools::load_all()
set_options()
```

```{r captions}
# Include captions below using `captioner` package via library(captioner)
stab <- captioner::captioner(prefix = 'ESM Table')
cite_st <- pryr::partial(stab, display = 'cite')

stab_qic <- stab('qic', 'Comparing generalized estimating equation models adjusting for different covariates using Quasi-Likelihood Information Criterion.')
stab_dist <- stab('nefa', 'Raw concentration values of each non-esterified fatty acid in the PROMISE cohort at the baseline visit (2004-2006). Values are presented as mean (SD).')
stab_gee_unadj <- stab('gee_unadj', 'Unadjusted generalized estimating equation models of the longitudinal associations of individual non-esterified fatty acids (mol% and nmol/mL) with insulin sensitivity and beta-cell function over the 6 years in the PROMISE cohort.  GEE models are only adjusted for time. Outcome variables were log-transformed, predictor variables were scaled, and x-axis values were exponentiated to represent percent difference per SD increase in the fatty acid.  P-values were adjusted for the BH false discovery rate, with significant (p<0.05) associations indicated by asterisk.')
stab_gee_adj <- stab('gee_adj', 'Longitudinal associations of individual non-esterified fatty acids (mol% and nmol/mL) with insulin sensitivity and beta-cell function over the 6 years in the PROMISE cohort. GEE models were adjusted for time, sex, ethnicity, baseline age, WC, ALT, and family history of diabetes. Outcome variables were log-transformed, predictor variables were scaled, and x-axis values were exponentiated to represent percent difference per SD increase in the fatty acid.  P-values were adjusted for the BH false discovery rate, with significant (p<0.05) associations indicated by asterisk (*).')
```

# ESM Tables

```{r ESM_Table1, results="asis"}
analyze_qic() %>% 
    table_qic(stab_qic)
```

Given the number of possible combinations of outcome and predictor variables,
only ISI and ISSI-2 with total non-esterified fatty acids (nmol/mL) were used to
compare various GEE models and to select a final model.  Baseline age was used
as including both the original age and the time variable would result in
collinearity.  Column names are: QIC is the quasi-likelihood information criteria
(smaller values, eg. larger negative values, indicate a better fit compared to
other models), Delta is the QIC minus the lowest QIC (models that have a delta <10 
between them are considered equivalent).  Models were:

- M0: log(ISSI-2) or log(ISI) = total non-esterified fatty acids (nmol/mL) + years from baseline
- M1: M0 + NEFA by time interaction
- M2: M0 + sex + ethnicity + baseline age
- M3: M2 + waist
- M4: M3 + ALT
- M5: M3 + physical activity
- M6: M3 + alcohol intake
- M7: M3 + family history of diabetes
- M8: M3 + smoking status
- M9: M3 + ALT + physical activity
- M10: M3 + NEFA by time interaction
- M11: M9 + NEFA by time interaction

```{r ESM_Table2}
table_distribution(project_data, stab_dist)
```

```{r ESM_Table3_4}
gee_results %>% 
    dplyr::filter(model == "Unadjusted") %>% 
    table_gee_main(stab_gee_unadj)
gee_results %>% 
    dplyr::filter(model == "Adjusted") %>% 
    table_gee_main(stab_gee_adj)
```